---
title: "Impact of atmospheric CO2 concentration on precipitation"
author: "Robert Wright"
date: "2023-01-11"
output:
  html_notebook: default
  github_document: default
  html_document:
    df_print: paged
  pdf_document: default
---

Investigating precipitation changes under different global warming scenarios (i.e., different CO2 concentrations).

```{r libraries, include=FALSE}
library("tidyverse")

knitr::opts_chunk$set(
  fig.width = 6, fig.asp = 0.618, out.width = "70%", fig.align = "center",
  echo = TRUE, collapse = TRUE, comment = "#>", fig.show = "hold"
)
```


```{r}
sessionInfo()
```

# Outline

1. **Climate simulations**: I'm using three PLASIM runs with CO2 concentrations of 610ppm, 790ppm & 1090ppm (the numbers come from SSPs, see section below). The default ("reference") value is 360ppm. The simulations start in 2000 and last for 30 years.
2. **Preprocessing** (using `cdo`-commands):
    - Compute climatology from *global* mean temperature for all simulations. Remove 29 Feb. Use final 10 years of simulation and compute 5-day arithmetic mean. Get climatology (values and standard deviation) by averaging years.
    - Similarly for precipitation, but keeping spatial dimensions. (Ist das richtig so? Also äquivalent zur Temperatur?) Save data.
    - Then, compute (time) average and anomaly. **Note**: Since $\overline{x_i} - \overline{y_i} = \overline{x_i - y_i}$, it doesn't matter in which order anomaly and average are computed.
    
3. **Analysis** in this notebook:
    - Visualize temperature climatology. Compute anomalies (subtract reference climatology, Fehlerfortpflanzung!). Get global surface temperature warming $\Delta T_{s}$ for each simulation by (time-)averaging climatology (std's mitteln?).
    - Plot map of mean precipitation anomaly.
    - Investigate region in particular? Something else?

# Shared Socioeconomic Pathways

The atmospheric CO2 concentration can be derived from these projected socioeconomic pathways, I use the preprocessed data from [ourworldindata](https://ourworldindata.org/explorers/ipcc-scenarios?facet=none&Metric=Greenhouse+gas+concentrations&Sub-metric=Carbon+dioxide+%28CO%E2%82%82%29&Rate=Per+capita&Region=Global&country=SSP1+-+Baseline~SSP2+-+Baseline~SSP3+-+Baseline~SSP4+-+Baseline~SSP5+-+Baseline).

```{r ssps import}
ssps <- read_csv(
  "../data/ipcc-scenarios.csv",
  col_select = c("Scenario", "Year", "CO2 concentration")
  ) %>% 
  rename(CO2_concentration=`CO2 concentration`) # remove whitespace in column label
```
```{r ssps plots}
ssps %>% 
  filter(grepl("Baseline", Scenario)) %>%  # keep "baseline" scenarios
  ggplot(aes(Year, CO2_concentration, color = Scenario)) +
    geom_line() +
    geom_point() +
    labs(title = "Shared Socioeconomic Pathways",
         x = "year",
         y = "CO2 concentration")
  #ggsave("../figs/ssps.png")
```

# 